---
title: "Week-07-Notes"
author: "Andrew Nalundasan"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(haven)
library(here)
library(readxl)
library(vtable)
library(corrplot)
library(GGally)
library(jtools)
library(gridExtra)
library(janitor)
```

# "Can Surveillance AI Make the Workplace Safe?" (Chamorro-Premuzic)

+ surveillance - ubiquitous temptation by people in power to exercise control

    + typically have a negative connotation
    + natural tendency is to push back on surveillance 
    
+ surveillance can monitor for COVID-19 and keep workforce safe & healthy
+ "creepy" is a common emotion when talking about surveillance

    + creepy <- potentially dangerous, not actually dangerous (yet)
    + leads to anxiety and nervousness

+ values discussion: 

    + autonomy (aka "freedom")
        + value doing things our own way on our own time
        + surveillance threatens autonomy
        + autonomy is important to people as a motivator
        + opposite of micromanaging
    + trust
        + surveillance done poorly is the opposite of trusting employees
        + the more i'm surveilled, the less I'm trusted
        + the less I'm trusted, the less you're trusted

+ digital lifestyles make surveillance easier to achieve
+ amazon - corporate surveillance in terms of bodily work

    + want to move to monitoring "intellectual activity" 
    + this is now possible with big data and digital lifestyles
    
+ legal constraints vs. ethical constraints

    + corporate surveillance is a major power imbalance 
    + ethics is about **CONSENT**
    
+ introduction of ethical constraints

    1. "no surprise rule" - organizations must make employees aware of what they're monitoring
        + knowing what data is collected and **WHY**
        + if an employer can communicate WHY they're collecting the data, employees are more likely to give buy-in
    2. employees must have option to opt in or out
        + opt-in in voluntarily 
        + there should be a benefit to data collection
        + people are more likely to opt-in if they understand how it benefits them
    3. There should be a real benefit to employees and not just the company
        + corporate surveillance:
            1. how you're going to do it
            2. how to explain it to the employees
            3. how employees will benefit by opting in
        
+ Analytics professionals: 
    
    + handle the data (wrangle)
    + interpret the data
    + communicate the data
    
# Puzzle of Motivation (TED Talk, Daniel Pink, July 2009, Oxford, England)

+ Don't want to tell a story, want to make a case on rethinking how to run our businesses
+ Overcome functional fixedness - use the container of the thumb tacks to use as a platform for the candle
+ Power of incentives - if you want people to perform better, reward them via incentives

    + incenctives dulls thinking and blocks creativity
    + contingent motivators ("if you do this, then you get that")
    + "There's a mismatch between what science knows and what business does"
    + higher incentives leads to worse performance
    + LSE: "financial incentives results in poor business performance"
    
+ intrinsic interests

    1. autonomy - urge to direct our own lives
    2. mastery - desire to get better at something that matters
    3. purpose - yearning to do what we do in the service of something larger than ourselves
    
+ **management** leads to compliance
+ **self direction** leads to engagement
+ ROWE - Results Only Working Environment
    
    + people don't have schedules. They have complete autonomy on how they get their work done
    
+ intrinsic motivators vs. extrinsic motivators

# "How to Monitor Employees - While Respecting Their Privacy" (Blackman)
Blackman, R. 2020, May. *How to monitor your employees - while respecting their privacy*. Harvard Business Review

+ Employee Monitoring <- Corporate Surveillance
+ Goal ("end") of corporations: increase employee productivity
    
    + maximize efficiency
    + don't want employees to take advantage of company

+ talking with c-level executives: need to talk in terms of "risk"

    + corporations don't care much for ethics. 
    + want to know how risk is being introduced to the company
    
+ 2 fundamental values: 

    1. Trust
    2. Autonomy - doing things your own way
    
+ micromanagement is the opposite of respecting employee's autonomy

+ 6 principles of respecting employees privacy while surveilling them: 

    1. Choose metrics carefully by involveing all relevant stakeholders
        + monitor only relevant and necessary data 
        + must be intentionaly with the metrics chosen to monitor productivity
    2. Be transparent about what's being monitored and why
        + take time to communicate openly and honestly
        + improved buy-in when communication takes place and not being secretive about surveillance
    3. Offer carrots as well as sticks
        + donkey is motivated by a stick (beat donkey with stick) or carrot (donkey follows the carrot)
        + don't just use monitoring as punitive measures, but make it an incentive system
    4. Good workers are rarely good workers all the time
        + People are not robots
        + Humans have lives and can ebb and flow in productivity
        + Robots are expected to have same level of productivity over time
    5. Make sure thate are no disproprotionate effects
        + Discriminatory monitoring - 400 years of white supremacy, 1000s of years of patriarchy
        + monitoring ends up being discriminatory and harming lower level groups
    6. Decrease monitoring when and where you can
        + There's no point in checking this, so stop monitoring specific areas
        
        
# "HR Analytics at ScaleneWorks" (Rahul and Kumar)

## Read in data and explore
        
```{r}
scalene_works <- read_csv(here("DataSet_HR Analytics at Scalene Works.csv"))
vtable(scalene_works)
summary(scalene_works)
head(scalene_works, 10)
```

## Rough visuals and exploration

```{r}
p <- scalene_works %>% 
  ggplot(mapping = aes(x = Age, fill = Status))

p + geom_histogram(bins = 10, position = "dodge") + 
  labs(title = "Employees joined by age")

```

**Comments**

+ Early 30s is biggest source of 'joined' employees
+ Early 30s has lowest 'renege' rate
+ Late 20s have slightly higher 'renege' rate than early 20s
+ Late 20s have slightly lower 'joined' rate than early 20s
+ Late 30s has 'renege' rate of ~ 50%

```{r}
p <- scalene_works %>% 
  ggplot(mapping = aes(x = Gender, fill = Status))

p + geom_histogram(bins = 10, stat = "count", position = "dodge") + 
  labs(title = "Employees joined by Gender")
```

**Comments**

+ Males have higher 'renege' rate, ~ 25%
+ Females 'renege' rate is almost 50%
+ Significantly more Males seem to be hired 

```{r}
p <- scalene_works %>% 
  ggplot(mapping = aes(x = `Candidate Source`, fill = Status))

p + geom_histogram(bins = 10, stat = "count", position = "dodge") + 
  labs(title = "Employees joined by Candidate Source")
```

**Comments**

+ Direct candidate resources yields the most amount of 'joined' employees as well as 'not joined'

    + Still, Direct is the best source for 'joined' 

+ Agency appears to have ~2:1 ratio of 'joined' vs. 'not joined'. 

    + Hires made via Agency seem to have the largest 'renege' rate
    
+ Employee Referral has smallest amount of 'not joined' employees

    + Employee Referral also is the smallest source of 'joined' employees
    
    
```{r}
p <- scalene_works %>% 
  ggplot(mapping = aes(x = `Joining Bonus`, fill = Status))

p + geom_histogram(bins = 10, stat = "count", position = "dodge") + 
  labs(title = "Employees joined by Joining Bonus")
```

**Comments**

+ Very surprising. 
+ Half of candidates that were offered signing bonuses ended up renegging 
+ 25% of candidates not offered a signing bonus ended up renegging
+ 8000+ candidates joined the company without being offered a signing bonus

```{r}
p <- scalene_works %>% 
  ggplot(mapping = aes(x = `Offered band`, fill = Status))

p + geom_histogram(bins = 10, stat = "count", position = "dodge") + 
  labs(title = "Employees joined by Offered Band")
```

**Comments**

+ 20% of E0s renege
+ 42% of E1s renege
+ 46% of E2s renege
+ 42% of E3s renege

```{r}
p <- scalene_works %>% 
  ggplot(mapping = aes(x = `DOJ Extended`, fill = Status))

p + geom_histogram(bins = 10, stat = "count", position = "dodge") + 
  labs(title = "Employees joined by Date of Joining Extension")
```

**Comments**

+ If a Date Of Joining was extended, likelihood of candidate joining is higher
+ Perhaps a driver of renegging is requesting an DOJ Extension to better weigh options. If DOJ Extension is not granted, candidate will just renege. 
+ If DOJ Extension is granted, renege rate decreases (Join rate also decreases, but doesn't seem to be practically significant)

```{r}
p <- scalene_works %>% 
  ggplot(mapping = aes(x = `Duration to accept offer`, fill = Status))

p + geom_histogram(stat = "count", position = "dodge") +
  scale_x_log10() +
  #scale_x_continuous(breaks = 25) +
  labs(title = "Employees joined by Duration to accept offer")
```


```{r}
# Only use bar graphs (like histogram) because nothing like boxplot for cateorical variables
grid.arrange(

# Status
ggplot(data = scalene_works, mapping = aes(x = Status)) +
  geom_bar(),

# Age
ggplot(data = scalene_works, mapping = aes(x = Age)) +
  geom_bar(), 

# Gender
ggplot(data = scalene_works, mapping = aes(x = Gender)) +
  geom_bar(),

# Joining Bonus
ggplot(data = scalene_works, mapping = aes(x = `Joining Bonus`)) + 
  geom_bar(),

ggplot(data = scalene_works, mapping = aes(x = `Candidate Source`)) + 
  geom_bar(),

ggplot(data = scalene_works, mapping = aes(x = Location)) + 
  geom_bar(),

#ggplot(data = scalene_works, mapping = aes(x = `Percent hike expected in CTC`)) + 
#  geom_bar(),

#ggplot(data = scalene_works, mapping = aes(x = `Percent hike offered in CTC`)) + 
#  geom_bar(),

#ggplot(data = scalene_works, mapping = aes(x = `Offered band`)) + 
#  geom_bar(),

#ggplot(data = scalene_works, mapping = aes(x = `Notice period`)) + 
#  geom_bar(),

#ggplot(data = scalene_works, mapping = aes(x = `Duration to accept offer`)) + 
#  geom_bar(),

ncol = 2 )
```

**Comments**

+ Outputs counts of each category of each variable

```{r}
# Proportion contingency/cross table
scalene_works %>% 
  tabyl(Status, Gender) %>% 
  adorn_totals(where = c("row", "col")) %>% 
  adorn_percentages(denominator = "all") %>% # creates proportions
  adorn_rounding(2) 
```

```{r}
scalene_works %>% 
  tabyl(Status, `Candidate Source`) %>% 
  adorn_totals(where = c("row", "col")) %>% 
  adorn_percentages(denominator = "all") %>% # creates proportions
  adorn_rounding(2)
```


```{r}
scalene_works %>% 
  tabyl(Status, `Offered band`) %>% 
  adorn_totals(where = c("row", "col")) %>% 
  adorn_percentages(denominator = "all") %>% # creates proportions
  adorn_rounding(2)
```

**Comments**

+ 1% of E0s renege
+ 17% of E1s renege
+ 9% of E2s renege
+ 2% of E3s renege

+ ~30% of candidates that accept the job offer end up renegging 

```{r}
factor_scalene <- scalene_works %>% 
  mutate(factor_DOJ = as.factor(`DOJ Extended`)) %>% 
  mutate(factor_band = as.factor(`Offered band`)) %>% 
  mutate(factor_bonus = as.factor(`Joining Bonus`)) %>% 
  mutate(factor_gender = as.factor(Gender)) %>% 
  mutate(factor_source = as.factor(`Candidate Source`)) %>% 
  mutate(factor_LOB = as.factor(LOB)) %>% 
  mutate(factor_location = as.factor(Location)) %>% 
  mutate(factor_status = as.factor(Status)) 

vtable(factor_scalene)

is.factor(factor_scalene$factor_DOJ)
contrasts(factor_scalene$factor_band)
```

## Model Fitting

```{r}
train_scalene <- factor_scalene[1:9866,] # 80% of dataset
test_scalene <- factor_scalene[9867:12333,] # 20% of dataset
```


```{r}
# fit the model
model <- glm(factor_status ~ ., family = binomial(link = 'logit'), data = train_scalene)
export_summs(model)
summary(model)
```


```{r}
scalene_corr <- scalene_works %>% 
  select_if(is.numeric) %>% # Use to select just the numeric variables
  cor() %>% 
  round(2) %>% 
  kable()

scalene_corr
```

**Comments**

+ Biggest positive correlations:

    + Age + Condidate Ref (0.36)
    + Notice Period + Duration to accept offer (0.30)
    + Rex in Yrs + Notice period (0.27)
    + Percent hike offered in CTC + Percent hike expected in CTC (0.64)
    + Percent difference CTC + Percent hike offered in CTC (0.57)
    + Age + Rex in Yrs (0.48)
    
+ Biggest negative correlations

    + Notice period + Candidate Ref (-0.22)
    + Percent difference CTC + Percent hike expected in CTC (-0.18)

```{r}
corr_sm <- scalene_works %>% select(c("Joining Bonus", "Gender", "Candidate Source", "Age", "Status"))

ggpairs(data = corr_sm, mapping = aes(color = Status),
        upper = list(continuous = wrap("density"), combo = "box_no_facet"),
        lower = list(continuous = wrap("points"), combo = wrap("dot_no_facet")))
```

**Comments**

+ I don't know how to read this. This doesn't give me much to work off of at all

### Regressions

+ Ho: some activity does NOT impact the variable of interest
+ Ha: some activity DOES impact the variable of interest

+ Variable of interest <- Status
+ Ho: Gender does not impact Status
+ Ha: Gender impacts Status

Data Transformations:

```{r}
num_scalene <- scalene_works %>% 
  mutate(num_status = recode(Status,
                                "Not Joined" = 0,
                                "Joined" = 1)) %>% 
  mutate(num_gender = recode(Gender,
                                "Female" = 0,
                                "Male" = 1)) %>% 
  mutate(num_join_bonus = recode(`Joining Bonus`,
                                 "No" = 0,
                                 "Yes" = 1))

```

Wilcox statistical test:

```{r}
wilcox.test(num_scalene$num_gender[num_scalene$num_status == 0], 
            num_scalene$num_gender[num_scalene$num_status == 1], conf.int = TRUE)
```

**Comments**

+ p-value is tiny, so reject the Ho. 
+ Gender does impact status!!

linear regression:

```{r}
reg_gender <- lm(num_status ~ num_gender, num_scalene)
export_summs(reg_gender)
```

**Comments**

+ gender has a slight negative impact on status

```{r}
reg_join_bonus <- lm(num_status ~ num_join_bonus, num_scalene)
export_summs(reg_join_bonus)
```

**Comments**

+ Being offered a joining bonus does impact status


